Do convolutional neural networks exhibit human-like attentional effects? Conflicting results from a visual search task

### 2019-05-28 docs/ccneuro2019/paper.tex
yet another rewrite of abstract in ccneuro/paper.tex

\section{Abstract}
{
\bf
Selective attention mechanisms preferentially route some sensory inputs for further 
processing, while necessarily discarding others. Researchers studying visual
attention have often divided selective attention into overt and covert mechanisms. 
Overt mechanisms involve orienting the eyes and/or head, while covert mechanisms 
occur "centrally", in the brain, without overt orientation. Cognitive scientists have long
used a new classic form of visual search task to investigate theoretical mechanisms of 
selective covert visual attention. Here we build on previous studies that examine the 
parallels between the visual system and convolutional neural networks (CNNs), by testing 
whether CNNs exhibit human-like attentional effects while performing this visual search 
task. Following a typical approach, we first show that CNNs trained 
for image classification can show human-like attentional effects. Then we demonstrate 
that, by changing how we train the networks and the statistics of the training data, 
these effects can largely be eliminated, i.e., that CNNs are capable of performing the 
task without explicit attention mechanisms. We go on to show that manipulating the 
discriminability of the target and distractors in the task can produce attentional 
effects in spite of the changes we made to training. Based on these results, we conclude 
that CNNs do not show attentional effects.
}
\begin{quote}
\small
\textbf{Keywords:} 
attention;visual search
\end{quote}

>>>>

\section{Abstract}
{
\bf
Convolutional neural networks (CNNs) trained to perform tasks such as image classification
with human-like accuracy acquire representations similar to those in the primate 
visual system. Do CNNs that are pre-trained to use such representations conversely exhibit 
human-like behavior? We test this with a visual search task that has long been used to 
investigate attention mechanisms. We first show that CNNs using weights trained for image 
classification can exhibit human-like attentional effects when performing this visual 
search task. However, we then demonstrate that these effects can be 
eliminated, without changing the lower-level learned representations, by modifying other 
parameters of training and the balance of classes in the training data. One explanation 
for this finding is that we are only training networks to discriminate single types of 
visual search stimuli. We further test CNNs trained to perform multiple tasks.
We go on to show that manipulating the 
discriminability of the target and distractors in the task can produce attentional 
effects in spite of the changes we made to training. Based on these results, we conclude 
that CNNs do not show attentional effects.
}
\begin{quote}
\small
\textbf{Keywords:} 
attention;visual search
\end{quote}
